HEP Software Foundation Community White Paper Working Group -- Training, Staffing and Careers

### 2018-06-13 CWP/papers/HSF-CWP-2017-02_training/latex/cwp-wg-training-careers.tex
Adding author/institute, and formatting (#85)
The following changes were made just to the author list:
* I added myself "Christopher Tunnell" and my institute "University of Chicago".
* Furthermore, there were no commas after a few other names toward the end that I added for consistency
* The "and" I assume happens with the last name in alphabetical order.  Since "Tunnell" is after "Stewart", I moved the 'and' to my name as the last alphabetically.
Mention IN2P3 gitlab set of tutorials.

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{xspace}

@@

\abstract{The rapid evolution of technology and the parallel increasing 
complexity of algorithmic analysis in HEP require developers 
to acquire a much larger portfolio of programming skills. Young 
researchers graduating from universities worldwide currently do not receive adequate 
preparation in the very diverse fields of modern computing to respond to growing needs 
of the most advanced experimental challenges. There is a growing cnsensus in the HEP community 
on the need for training programmes to bring researchers up to date with new software technologies, 
in particular in the domains of concurrent programming and artificial intelligence. We review some
of the initiatives under way for introducing new training programmes and highlight some of the issues 
that need to be taken into account for these to be successful.}

@@

\begin{tabular*}{\linewidth}{lc@{\extracolsep{\fill}}r@{\extracolsep{0pt}}}
 & & HSF-CWP-2017-02 \\
 & & June 6, 2018 \\ % use \date or hardwire e.g. December 15, 2017
 & & \\
\end{tabular*}
\vspace{2.0cm}

@@

\tableofcontents
\clearpage

@@

The field of high energy physics (HEP) has consistently exploited 
the latest innovations in computational tools and technologies for processing data.
The software toolset of the particle physicist is ever-growing, and the problem
set increasingly complex. Young researchers will likely encounter a number of 
programming techniques with which they are totally unfamiliar, and will therefore 
need "on-the-job" training in order for them to be productive.

A strong research community comprises independent thinkers who ask
research questions and direct their own analyses, instead of blindly following
prescriptions and recipes. It is therefore imperative that training opportunities 
and resources are made available, such that adequate career
path possibilities exist for people in HEP who would otherwise leave the field
due to limited advancement opportunities.

@@

%____________________________________________________________________
%____________________________________________________________________
\section{Motivation for some HSF coordinated actions}

@@

As tools become more complex, physicists must be continuously retrained in order
to utilize them effectively. It is important to find a way to deploy training efforts
holistically and cross-experimentally, whenever possible, in order to avoid unnecessary 
duplication of the effort involved. This will require organization and coordination to
ensure that different software teams can agree on common standards, thereby allowing different
experiments to adopt the same olutions in an efficient way.

Beyond the immediate benefit to the scientific community in having well-trained
collaborators, most people who start a physics graduate programme will have
careers outside of academia. This trend is reflected in several calls for
proposals from national scientific funding agencies that promote training in computer
infrastructure areas, such as the EU's Horizon 2020 programme and those from
the National Science Foundation (NSF) [Refs 1 and 2].

An important challenge will be to discover incentives that will encourage 
increasing numbers of people to dedicate time and effort to train their colleagues.
To be effective, the training must be done by people who work at the cutting
edge of the technology, who in many cases are found among the youngest
researchers. These are also the very same people who are most in need of
officially recognized credits for the advancement of their careers. Currently
visibility and recognition is given only to those working on a data analysis project, 
rather othan on the development and support of the underlying software. A possible 
solution to this conundrum could be the establishment of specific career paths 
for researchers specialising in scientific software development. However, the practical 
implementation of such an approach is extremely challenging, as there are a wide range of
institutions involved that belong to different countries with their own
policies, priorities, and funding strategies.

Broadly speaking, two scientific profiles of researchers could be
envisaged; a physicst could make a detailed plan of what they need, expressed 
in a tidy requirements document, and a computer scientist could use this document 
to provide the required solution. This is how the software development process 
typically works in the software industry. However, in HEP requirements are rapidly
changing forcing developers and physicsts to closely interact during the entire process of
software development; this works smoothly only when both communities have the
same goals and speak the same language, hence the need for physicists with a good
knowledge of computer science.

@@

Both career paths suffer in academic terms from few or poorly cited
publications, few opportunities to win grants in their own right and little
opportunity for impact with commerce and wider society. These three failings
make an academic appointment challenging or disfavoured compared to other
applicants. However, there are measures that an be taken to address all three
weaknesses.

@@

For purposes of training, it is useful to think of all our user community as
\emph{students}. More specifically, we can broadly identify four different sets
of students, classified by experience level, recognizing that even within these
groups there will be people who will learn at different speeds and with
differing depths of understanding:

\begin{itemize}
   \item \emph{Beginners}: New collaborators with no knowledge of the 
   tools or techniques they are expected to use. These might be undergraduates 
   but they could also be senior researchers who have not yet or not recently 
   engaged in software analysis. These are people in need of some kind of formal 
   training in modern computing techniques (basically compiled and scripting 
   languages together with operating system basics such as filesystems, version 
   control, and command-line shells).
   \item \emph{Intermediate level analysts}: people with some experience in 
   either analysis concepts and/or tools, but  looking to supplement their 
   experience with more recent and modern approaches. 
   \item \emph{Advanced level analysts}: experts who have mastered current 
   technologies and implementations and who want to stay up-to-date with 
   new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software 
   development in areas not limited only to analysis, such as DAQ systems, 
   computing infrastructure, databases, pattern recongition, and complete 
   frameworks.
\end{itemize}

Each of these groups has very different training needs. However,
\emph{whenever possible}, any training programme should take advantage of
developments in pedagogy, such as active
learning\footnote{\url{http://www.crlt.umich.edu/tstrategies/tsal}}, peer
learning\footnote{\url{https://en.wikipedia.org/wiki/Peer_learning}}, and
web-based training.\footnote{See \S\ref{sec:initiatives}.} In some cases, it may
even be advantageous to hand out code samples that are purposely broken or
flawed, and ask students to fix or improve them. Learning the material in a way
that sticks is difficult and challenging for both the students
and the instructor and often takes more time than we would prefer. However, this
is the best way to educate scientists who can fully
contribute to the physics programmes at large, which is really the ultimate goal
of any training programme.

@@

At all stages of software and computing training, we should take care to
encourage \emph{Good Practices Across the Community} (GPAC), such as error
checking, modularity of code design, version control, writing tests, etc. All
the key concepts addressed in the training should not be specific to a
particular experiment or field of application, but general enough to be useful
to the whole HEP community and possibly beyond. In this section, we present a
list of specific concepts that need to be taught, in
order to guarantee the basic level of competence needed to write efficient code
for the different tasks performed in HEP experiments.

An essential knowledge base that needs to be transferred includes basic
programming concepts, data structures, basics of code design, error checking,
code management tools, validation and debugging tools. More advanced topics
include modularity of code design, advanced data structures, evaluation metrics,
writing tests and working with different types of hardware accelerator.
Cutting-edge approaches such as the use of hardware accelerators (whether GPUs
or FPGAs), Deep Neural Networks, and Machine Learning should be considered
important training topics. Additionally, special emphasis should be given to
reporting results and documenting them.
The training subjects considered important to pursue are
listed as follows:

%____________________________________________________________________
%____________________________________________________________________
\section{Implementation Roadmap}
The implementation of training should target multiple training formats and knowledge transfer, such as videos, wikis, lectures, jupyter notebooks and advanced visualizations, etc.
People who learn in non-conventional ways should not be excluded from the
opportunity to learn. Since there are different stages of knowledge transfer,
programmes should be developed that encourage experts to share knowledge. Issues
considered in this section include global training initiatives, strategies for
leveraging existing forums, and training tool development.

@@

MOOCs can be used to develop an open-source set of tutorials and tools. Existing 
online courses such as Udacity and Coursera can be evaluated and recommended to 
the community. A novel approach such as WikiToLearn
could also explored to assess potential benefits, as has already been attempted
in the context of the GridKa School of Computing (see below).

@@

Question-and-answer websites, such as Stack Overflow for HEP, are also a very
useful resource for addressing common problems and questions. This approach has already
been considered by HSF start-up members, and turned out to be difficult to
pursue for a number of reasons, but in the future, boundary conditions might
change, making this approach viable.

@@

Creating an expert tutor volunteer network is another way to provide training
and support to the community.
This of course requires proper recognition, at least in terms of career
prospects as an incentive for young researchers. It is clear that the best
possible tutors are, in principle, those people who are actively engaged in
modern software developments: these are in general young researchers in the
first steps of their careers and, in order to be attracted to devoting
substantial time to training and tutoring, they must be assured that such an
effort would be properly recognised in an official way. Such kind of recognition
is not currently implemented, at least not in a standardized or official way. A
possible structure for such a network could be the establishment of a {\em
federation} of existing schools.

It should be considered that some professors, who act as graduate student
advisors, might need to be encouraged to make sure their students are properly
trained. Sometimes, students are instead pushed to learn the bare minimum to get
the work done, at the expense of a broader training/education curriculum that
could actually yield improved results further down the line. One incentive
would be to provide training programmes that also count as course credit, perhaps
as an elective. This model is already in limited use with some online
solutions\footnote{\url{https://www.edx.org/credit}}, but this is not universal.
Discussions should be started with collaborators at higher education
institutions to see what the roadblocks or opportunities would be for these
training sessions to serve double duty.
It should also be considered that not all students will end their career in
research or academia: their contribution to the research activity, as students,
should therefore also provide them knowledge, know-how, and skills considered a
valuable asset by industry, in order to increase their chances of a career
outside research.

@@

efforts such as the CERN School of Computing's Bertinoro and Kit ones): instead

@@

   \item Their material, made public on a web-site, can be further expanded by collaborators
   (also at their own pace) or, even better and more productively, by students

@@

An example of a such a web-based platform already exists, and has been
implemented as an Open Source project (backed by Wikimedia) by a group of more
than 30 Italian students. The project is WikiToLearn
\footnote{\url{https://it.wikitolearn.org/}}. It hosts training material in
several languages, for several disciplines, ranging from Economics to Physics,
Mathematics, and several others. Because it is based upon
wikimedia\footnote{\url{https://www.wikimedia.org/}} software, it is very easy
to add material to the site, and to make it appear under a specific topic (such
as Software/Techniques/Machine-Learning) and to manipulate it as if it
were a single document. In the end, students can selectively choose individual
chapters from the site and have the corresponding pdf sent them as a book,
complete with index, content, and chapters.

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the learning curve has been kept appropriately shallow.

An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL\footnote{\url{https://en.wikitolearn.org/GridKa2017}}.
This constitutes an interesting example of what can be accomplished using this
platform; it is just a first example of what's possible, but an inspiring one.

Finally it should be important to evaluate, if and to which extent, the
complementary approaches to training (such as schools and dedicated web-site
courses) could be of mutual benefit in the development of a complete training 
programme.

@@

Within the HEP community, there are already some working examples of dedicated
training environments that alternate between general topics and
experiment-specific topics. The LHC Physics Center (LPC) at Fermilab hosts
Hands-on Advanced Tutorial Sessions
(HATS)\footnote{\url{http://lpc.fnal.gov/programs/schools-workshops/}}
throughout the year to introduce and train participants in topics as diverse as
the latest $b$-tagging algorithms, Git/GitHub, and machine learning. These HATS
provide face-to-face time with instructors and participants at Fermilab, and
also allow remote collaborators to join in by video and complete the same online
exercises. A similar approach is in use in the CMS Data Analysis Schools
(CMSDAS)\footnote{\url{http://lpc.fnal.gov/programs/schools-workshops/cmsdas.shtml}},
a series of week-long workshops that now take place at multiple labs all over
the world and are designed to ramp up new collaborators in CMS-specific analysis
tools while providing some discussion of the physics as well.

Other examples are CERN School of
Computing\footnote{\url{https://csc.web.cern.ch/}}, CERN OpenLab Software
workshops\footnote{\url{http://openlab.cern/}} education in collaboration with
industry partners, and a series of more advanced topical training courses
provided by MPI Munich and DESY that focus on advanced programming, use of
acceleration hardware and statistical tools including machine learning. This
list includes the
Bertinoro\footnote{\url{https://web.infn.it/esc17/index.php?lang=en}}, 
GridKa\footnote{\url{http://gridka-school.scc.kit.edu/2017/}}, and
CODAS-HEP\footnote{\url{http://codas-hep.org}} Schools of Computing.

@@

There are a growing set of MOOCs teaching various subjects. Since there are many
options, there is a wide variety with respect to the depth of the material and
specific tools taught. Exploring these options allows us to choose which is the
right offering for the knowledge needed to work on a specific experiment. We
can pick and choose modules to tailor an appropriate roadmap of skills to learn.
More difficult will be to assemble specific training material not already
avaible elsewhere in an efficient and organized way, since this requires
adequate organization, volounteers and and a suitable infrastructure.

Several industry conferences already exist that bring together those in academia
and industry who are at the cutting edge of these techniques. Conferences such
as NIPS\footnote{\url{https://nips.cc}} and
PyData\footnote{\url{https://pydata.org}} provide a focused place where
attendees can learn a lot about machine learning in a short period of time.
Machine learning concepts such as current methods, tools, and problems facing
industry and academia can be learned at conferences. In addition, conferences
are an excellent networking opportunity; attendees can meet and share ideas with
fellow learners and experts. Bonds can be formed quickly at conferences that can
be maintained after the duration of the conference. These connections to the
outside community can be essential since we will be evolving training materials
to ensure that they stay relevant over time.

For example, at the time of this writing,
Coursera\footnote{\url{https://www.coursera.org/learn/machine-learning}} and
Udacity\footnote{\url{https://www.udacity.com/course/intro-to-machine-learning--ud120}}
both provide great machine learning massive open online courses at no cost.
These two courses both provide a great foundation for learning machine learning
fundamentals. However, Coursera's approach emphasizes more theory (with more
math background necessary) and uses MATLAB/Octave while Udacity's approach
emphasizes practical using machine learning techniques in a practical sense
using Python tooling.

@@

Training is something that a large cross-section of the community
understands to be important, but finding time and effort to contribute to this
effort is not actively on the radar of most potential contributors. Providing
incentives for their participation and creating the appropriate platforms can go
a long way to reach a productive training environment. It can be as simple as
inviting someone to give a software tutorial on the subject that they are
familiar with, give a lecture or seminar or contribute to a growing knowledge
base.

@@

It is also critical to incorporate training into grant proposals so that it can
be connected with other areas such as research and development. Efforts like
DIANA-HEP \footnote{\url{http://diana-hep.org/}} and aMVA4NewPhysics
\footnote{\url{https://amva4newphysics.wordpress.com/}} that combine training
and software development are good examples of such ideas in practice. More
examples of such efforts are needed.

@@

Software Carpentry (SC) \footnote{\url{https://software-carpentry.org}} and Data
Carpentry (DC) \footnote{\url{http://www.datacarpentry.org}} are two parts of The
Carpentries organization that collaboratively build and teach some of the basic
concepts in developing and maintaining software, and analyzing data,
respectively.  The materials that they develop and use are open, and can be
customized for science domains (e.g., HEP), or participant groups (e.g.,
undergraduates). Their model is that they offer training of trainers, and then
the trainers who have graduated can offer training under the SC/DC names, though
of course, anyone can use the SC/DC materials without doing it under the SC/DC
names.
Subjects covered in a typical Data Carpentry school are given below. This list
is taken from the curriculum of the CODATA-RDA Research Data Science Summer
School in progress in Trieste, Italy during July 2017.
(http://indico.ictp.it/event/7974/):

@@

%____________________________________________________________________
%____________________________________________________________________
%\begin{itemize}
%
%   \item How to motivate the community to participate in developing 
programs? Can some of it come from \& leadership?
%   \item How to deal with competitiveness in the field? Can’t force 
people to be interested in forming a community. How can we incentivize this? 
Building a %vibrant, welcoming community may help.
%\end{itemize}

@@

\printbibliography[title={references}]

>>>>

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage[titletoc]{appendix}

@@

\abstract{The rapid evolution of technology and the parallel increasing 
complexity of algorithmic analysis in HEP requires developers 
to acquire a much larger portfolio of programming skills. Young 
researchers graduating from universities worldwide currently do not receive adequate 
preparation in the very diverse fields of modern computing to respond to growing needs 
of the most advanced experimental challenges. There is a growing consensus in
the HEP community on the need for training programmes to bring researchers up to date with new software technologies, 
in particular in the domains of concurrent programming and artificial intelligence. We review some
of the initiatives under way for introducing new training programmes and highlight some of the issues 
that need to be taken into account for these to be successful.}

@@

\begin{tabular*}{\linewidth}{lc@{\extracolsep{\fill}}r@{\extracolsep{0pt}}}
 & & HSF-CWP-2017-02 \\
 & & June 20, 2018 \\ % use \date or hardwire e.g. December 15, 2017
 & & \\
\end{tabular*}
\vspace{2.0cm}

@@

\clearpage

@@

The field of high energy physics (HEP) has consistently exploited 
the latest innovations in computational tools and technologies for processing data.
The software toolset of the particle physicist is ever-growing, and the problem
set increasingly complex. Young researchers will likely encounter a number of 
programming techniques with which they are totally unfamiliar, and will therefore 
need ``on-the-job'' training in order for them to be productive.

Evidently a strong research community comprises independent thinkers who ask
research questions and direct their own analyses, instead of blindly following
prescriptions and recipes. It is therefore imperative that training opportunities 
and resources are made available, such that adequate career
path possibilities exist for people in HEP who would otherwise leave the field
due to limited advancement opportunities.

@@

%____________________________________________________________________
%____________________________________________________________________
\section{Career Paths for Software Experts}

@@

As tools become more complex, physicists must be continuously retrained in order
to utilize them effectively. It is important to find a way to deploy training efforts
holistically and cross-experimentally, whenever possible, in order to avoid unnecessary 
duplication of the effort involved. This will require organization and coordination to
ensure that different software teams can agree on common standards, thereby allowing different
experiments to adopt the same solutions in an efficient way.

Beyond the immediate benefit to the scientific community in having well-trained
collaborators, most people who start a physics graduate programme will have
careers outside of academia. This trend is reflected in several calls for
proposals from national scientific funding agencies that promote training in computer
infrastructure areas, such as the EU's Horizon 2020 programme and those from
the National Science Foundation (NSF) \cite{NSF1,NSF2}.

A key goal is to find incentives that will encourage 
increasing numbers of people to dedicate their time and effort to train their
colleagues.

To be effective, the training must be done by people who work at the cutting
edge of the technology, who in many cases are found among the youngest
researchers. These are also the very same people who are most in need of
officially recognized credits for the advancement of their careers. Currently
visibility and recognition is given mainly to those working on data analysis
projects, rather than to the development and support of the underlying
software. One solution to this problem would be the establishment of specific
career paths for researchers specialising in scientific software development. 
%% GS Comment - is this actually ``motivation'' - it seems more like a solution
% See suggested title change for this section...
However, the practical 
implementation of such an approach is extremely challenging, as there are a wide range of
institutions involved that belong to different countries with their own
policies, priorities, and funding strategies.

Broadly speaking, two scientific profiles of researchers could be
envisaged; a physicst could make a detailed plan of what they need, expressed 
in a tidy requirements document, and a computer scientist could use this document 
to provide the required solution. This is how the software development process 
typically works in the software industry. However, in HEP requirements are rapidly
changing, forcing developers and physicsts to interact closely during the entire
process of software development; this works smoothly only when both communities have the
same goals and speak the same language, hence the need for physicists with a good
knowledge of computer science.

@@

Both career paths suffer in academic terms from few or poorly cited
publications, few opportunities to win grants in their own right and little
opportunity for impact with commerce and wider society. These three failings
make an academic appointment challenging or disfavoured compared to other
applicants. However, there are measures that can be taken to address all three
weaknesses.
%% GS Comment - This is an odd place to end a section. What are these measures?
% Where do we list them? Even a forward reference would help.

@@

For purposes of training, it is useful to think of all our user community as
\emph{students}. More specifically, we can broadly identify four different sets
of students, classified by experience level:

\begin{itemize}
   \item \emph{Beginners}: New collaborators with no knowledge of the 
   tools or techniques they are expected to use. These might be undergraduates, 
   but they could also be senior researchers who have not yet, or not recently, 
   engaged in software analysis. These are people in need of some kind of formal 
   training in modern computing techniques (basically compiled and scripting 
   languages, together with operating system basics such as filesystems, version 
   control, and command-line shells).
   \item \emph{Intermediate level analysts}: people with some experience in 
   either analysis concepts and/or tools, but  looking to supplement their 
   experience with more recent and modern approaches. 
   \item \emph{Advanced level analysts}: experts who have mastered current 
   technologies and implementations and who want to stay up-to-date with 
   new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software 
   development in areas not limited only to analysis, such as DAQ systems, 
   computing infrastructure, databases, pattern recongition, and complete 
   frameworks.
\end{itemize}

Each of these groups has very different training needs. However,
\emph{whenever possible}, any training programme should take advantage of
developments in pedagogy, such as active
learning~\cite{ActiveLearning}, peer learning~\cite{PeerLearning}, and
web-based training.\footnote{See \S\ref{sec:initiatives}.} In some cases, it may
even be advantageous to hand out code samples that are purposely broken or
flawed, and ask students to fix or improve them. Learning the material in a way
that sticks is difficult and challenging for both the students
and the instructor and often takes more time than we would prefer. However, this
is the best way to educate scientists who can fully
contribute to the physics programmes at large, which is really the ultimate goal
of any training programme.

@@

At all stages of software and computing training, we should take care to
encourage \emph{Good Practices Across the Community}, such as error
checking, modularity of code design, version control, writing tests, etc. All
the key concepts addressed in the training should not be specific to a
particular experiment or field of application, but general enough to be useful
to the whole HEP community and beyond. In this section, we present a
list of specific concepts that need to be taught, in
order to guarantee the basic level of competence needed to write efficient code
for the different tasks performed in HEP experiments.

Essential knowledge that needs to be transferred includes basic
programming concepts, data structures, basics of code design, error checking,
code management tools, validation and debugging tools. More advanced topics
include modularity of code design, advanced data structures, evaluation metrics,
writing tests and working with different types of hardware accelerator.
Additionally, special emphasis should be given to
reporting results and documenting them.

Some of the training subjects considered important to pursue are
listed in Appendix \ref{appendix-a---training-topics-of-interest}.

% ____________________________________________________________________
% ____________________________________________________________________
\section{Implementation Roadmap}
The implementation of training should target multiple training formats and
knowledge transfer, such as videos, wikis, lectures, jupyter notebooks and
advanced visualizations, etc.
People who learn in non-conventional ways should not be excluded from the
opportunity to learn. 
%% GS Comment - I'm not really sure what that previous sentence is supposed to
% mean.
Since there are different stages of knowledge transfer,
programmes should be developed that encourage experts to share knowledge. Issues
considered in this section include global training initiatives, strategies for
leveraging existing forums, and training tool development.

@@

MOOCs can be used to develop an open-source set of
tutorials and tools. Existing online courses such as Udacity~\cite{Udacity} and 
Coursera~\cite{Coursera} can be
evaluated and recommended to the community. A novel approach such as WikiToLearn~\cite{WikiToLearn}
could also be explored to assess potential benefits (as has already been attempted
in the context of the GridKa School of Computing~\cite{GridKa}, see below).

@@

Question-and-answer websites such as Stack Overflow~\cite{Stackoverflow} for HEP are also a very
useful resource for common problems and questions. This approach has already
been considered by HSF start-up members, and turned out to be difficult to
pursue for a number of reasons, but in the future, boundary conditions might
change, making this approach viable.

@@

Creating an expert tutor volunteer network is another way to provide training
and support to the community.
This of course requires proper recognition, at least in terms of career
prospects as an incentive for young researchers. It is clear that the best
possible tutors are, in principle, those people who are actively engaged in
modern software developments: these are often young researchers in the
first steps of their careers and, in order to be attracted to devoting
substantial time to training and tutoring, they must be assured that such an
effort would be properly recognised in an official way. Such kind of recognition
is not currently implemented, at least not in a standardized or official way. A
possible structure for such a network could be the establishment of a 
{\em federation} of existing schools.

It should be considered that some professors, who act as graduate student
advisors, might need to be encouraged to make sure their students are properly
trained. Sometimes, students are instead pushed to learn the bare minimum to get
the work done, at the expense of a broader training/education curriculum that
could actually yield improved results further down the line. One incentive
would be to provide training programmes that also count as course credit, perhaps
as an elective. This model is already in limited use with some online
solutions~\cite{TrainingProgram}, but this is not universal.
Discussions should be started with collaborators at higher education
institutions to see what the roadblocks or opportunities would be for these
training sessions to serve double duty.
It should also be considered that not all students will end their career in
research or academia: their contribution to the research activity, as students,
should therefore also provide them knowledge, know-how, and skills considered a
valuable asset by industry, in order to increase their chances of a career
outside research.

@@

efforts such as the CERN School of Computing's Bertinoro and KIT ones): instead

@@

   \item Their material, available public on a web-site, can be further expanded
   by collaborators (also at their own pace) or, even better and more productively, by students

@@

An example of a such a web-based platform already exists, and has been
implemented as an Open Source project (backed by Wikimedia) by a group of more
than 30 Italian students. The project is WikiToLearn~\cite{WikiToLearn}. It hosts training material in
several languages, for several disciplines, ranging from Economics to Physics,
Mathematics, and several others. Because it is based upon
wikimedia~\cite{WikiMedia} software, it is very easy
to add material to the site, and to make it appear under a specific topic (such
as Software/Techniques/Machine-Learning) and to manipulate it as if it
were a single document. In the end, students can selectively choose individual
chapters from the site and have the corresponding pdf sent them as a book,
complete with index, content, and chapters.

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the necessary learning curve has been kept appropriately shallow.
An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL~\cite{WikiToLearnGridka}.
This constitutes an intersting example of what can be accomplished using this
the necessary learning curve has been kept appropriately shallow. An interesting exercise in
this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL\footnote{\url{https://en.wikitolearn.org/GridKa2017}}.
This constitutes an interesting example of what can be accomplished using this
platform; it is just a first example of what's possible, but an inspiring one.

Another example of web-based platform is a collection of online tutorials
(mostly written in french) hosted on the Gitlab IN2P3
server{\url{https://gitlab.in2p3.fr/MaitresNageurs}}. Taking profit of
the Gitlab ability to host Docker images, those tutorials aims to avoid the
"damned installation step" that often absorbs half of a training session.
Similarly to the GridKa School for WikiToLearn, the annual IN2P3 Computing
Days are an opportunity to refresh and extend the collection of
tutorials every year. Future work will focus on english translation, and the
development of a web site which will index the above tutorials, together with
the best recommended external tutorials.

Finally it should be important to evaluate, if and to which extent, the
complementary approaches to training, such as schools and dedicated web-site
courses, could be of mutual benefit, in other words how to make them efficiently cooperate in the
development of a complete training program.

@@

Within the HEP community, there are already some working examples of dedicated
training environments that alternate between general topics and
experiment-specific topics. The LHC Physics Center (LPC) at Fermilab hosts
Hands-on Advanced Tutorial Sessions
(HATS)~\cite{LPCHandsOn}
throughout the year to introduce and train participants in topics as diverse as
the latest $b$-tagging algorithms, Git/GitHub, and machine learning. These HATS
provide face-to-face time with instructors and participants at Fermilab, and
also allow remote collaborators to join in by video and complete the same online
exercises. A similar approach is in use in the CMS Data Analysis Schools
(CMSDAS)~\cite{LPCDataAnalysis},
a series of week-long workshops that now take place at multiple labs all over
the world and are designed to ramp up new collaborators in CMS-specific analysis
tools while providing some discussion of the physics as well.

Other examples are CERN School of Computing~\cite{CERNSchoolOfComputing}, 
CERN OpenLab Software workshops~\cite{OPENLab} education in collaboration with
industry partners, and a series of more advanced topical training courses
provided by MPI Munich and DESY that focus on advanced programming, use of
acceleration hardware and statistical tools including machine learning. This
list includes the
Bertinoro~\cite{Bertinoro}, 
GridKa~\cite{GridKa}, and
CODAS-HEP~\cite{CODAS-HEP} Schools of Computing.

@@

There are a growing set of MOOCs teaching various subjects. Since there are many
options, there is a wide variety with respect to the depth of the material and
specific tools taught. Exploring these options allows us to choose which is the
right offering for the knowledge needed to work on a specific experiment. We
can pick and choose modules to tailor an appropriate roadmap of skills to learn.
More difficult will be to assemble specific training material not already
avaible elsewhere in an efficient and organized way, since this requires
adequate organization, volounteers and a suitable infrastructure.

Several industry conferences already exist that bring together those in academia
and industry who are at the cutting edge of these techniques. Conferences such
as NIPS~\cite{NIPS} and
PyData~\cite{PyData} provide a focused place where
attendees can learn a lot about machine learning in a short period of time.
Machine learning concepts such as current methods, tools, and problems facing
industry and academia can be learned at conferences. In addition, conferences
are an excellent networking opportunity; attendees can meet and share ideas with
fellow learners and experts. Bonds can be formed quickly at conferences that can
be maintained after the duration of the conference. These connections to the
outside community can be essential since we will be evolving training materials
to ensure that they stay relevant over time.

For example, at the time of this writing,
Coursera~\cite{Coursera} and
Udacity~\cite{Udacity}
both provide great machine learning massive open online courses at no cost.
These two courses both provide a great foundation for assimilating machine learning
fundamentals. However, Coursera's approach emphasizes more theory (with more
math background necessary) and uses MATLAB/Octave while Udacity's approach
emphasizes more practical aspects using machine learning techniques 
using Python tools.

@@

Training is something that a large cross-section of the community
understands to be important, but finding time and effort to contribute to this
project is not actively on the radar of most potential volunteers. Providing
incentives for their participation and creating the appropriate platforms can go
a long way to reach a productive training environment. It can be as simple as
inviting someone to give a software tutorial on the subject that they are
familiar with, give a lecture or seminar or contribute to a growing knowledge
base.

@@

It is also critical to incorporate training into grant proposals so that it can
be connected with other areas such as research and development. Efforts like
DIANA-HEP~\cite{DIANA-HEP} and AMVA4NewPhysics~\cite{AMVA4NewPhysics}
that combine training
and software development are good examples of such ideas in practice. More
examples of such efforts are needed.

@@

Software Carpentry~\cite{SoftwareCarpentry} and Data
Carpentry~\cite{DataCarpentry} are two parts of The
Carpentries organization that collaboratively build and teach some of the basic
concepts in developing and maintaining software, and analyzing data,
respectively.  The materials that they develop and use are open, and can be
customized for science domains (e.g., HEP), or participant groups (e.g.,
undergraduates). Their model is that they offer training of trainers, and then
the trainers who have graduated can offer training under the SC/DC names, though
of course, anyone can use the SC/DC materials without doing it under the SC/DC
names.
Subjects covered in a typical Data Carpentry school are given in Appendix
\ref{appendix-b}.

@@

%% GS Comment - this section is incomplete. FIXME.

\newpage
\begin{appendices}

\sloppy
\raggedright

\hypertarget{appendix-a---training-topics-of-interest}{%
\section{Training Topics of Interest}\label{appendix-a---training-topics-of-interest}}

\hypertarget{appendix-b}{%
\section{Research Data Science Curriculum}\label{appendix-b}}

This list is taken from the curriculum of the CODATA-RDA Research Data
Science Summer School in progress in Trieste, Italy during July 2017.
(http://indico.ictp.it/event/7974/):

\end{appendices}

@@
